---
title: "HW 05 - Legos"
output: 
  tufte::tufte_html:
    css: hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  out.width = "80%",
  fig.asp = 0.618,
  fig.width = 10,
  dpi = 300
)
```

```{r load-packages, message = FALSE, eval = TRUE}
library(tidyverse)
library(dsbox)
```

1.  What are the three most common first names of purchasers?
```{r echo=TRUE}
lego_sales %>%
  count(first_name, sort = TRUE)
```
✅ In this sample, the three common first names of purchasers are *Jackson, Jacob, Joseph*

2.  What are the three most common themes of Lego sets purchased?
```{r echo=TRUE}
lego_sales %>%
  count(theme, sort = TRUE)
```

✅  In this sample, the three most common themes of Lego sets purchased are *Star Wars, Nexo Knights, Gear, and Mixels*

3.  Among the most common theme of Lego sets purchased, what is the most common subtheme?
```{r echo=TRUE}
lego_sales %>%
  filter(theme == "Star Wars") %>%
  count(subtheme, sort = TRUE)
```

✅  In this sample, the most common theme of Lego sets purchased is *Star Wars*, and the most common subtheme is *The Force Awakens*

4.  Create a new variable called `age_group` and group the ages into the following categories: "18 and under", "19 - 25", "26 - 35", "36 - 50", "51 and over".

```{r echo=TRUE}
lego_sales <- lego_sales %>%
  mutate(age_group = case_when(
    age <= 18 ~ "18 and under",
    age >= 19 & age <= 25 ~ "19 - 25",
    age >= 26 & age <= 35 ~ "26 - 35",
    age >= 36 & age <= 50 ~ "36 - 50",
    age >= 51 ~ "51 and over"
  ))
```

5.  Which age group has purchased the highest number of Lego sets.

```{r echo=TRUE}
lego_sales %>%
  count(age_group, sort = TRUE)
```

✅  In this sample group of *36-50* yo purchased the highest number of Lego sets.

6.  Which age group has spent the most money on Legos?

```{r echo=TRUE}
lego_sales %>%
  mutate(
    amount_spent = us_price * quantity
  ) %>%
  group_by(age_group) %>%
  summarise(
    total_spent = sum(amount_spent)
  ) %>%
  arrange(desc(total_spent))
```

✅  In this sample group of *36-50* yo has spent the most money on Legos.

7.  Which Lego theme has made the most money for Lego?

```{r echo=TRUE}
lego_sales %>%
  mutate(
    amount_spent = us_price * quantity
  ) %>%
  group_by(theme) %>%
  summarise(
    total_spent = sum(amount_spent)
  ) %>%
  arrange(desc(total_spent))
```

✅  In this sample *Star Wars* has made the most money for Lego

8.  Which area code has spent the most money on Legos?
    In the US the area code is the first 3 digits of a phone number.


```{r echo=TRUE}
lego_sales %>%
  mutate(
    amount_spent = us_price * quantity
  ) %>%
  group_by(substr(phone_number, 1, 3)) %>%
  summarise(
    total_spent = sum(amount_spent)
  ) %>%
  arrange(desc(total_spent))  
```  

✅ In this sample area code of *956 (state of Texas)* has spent the most money on Legos
    

9.  Come up with a question you want to answer using these data, and write it down.
    Then, create a data visualization that answers the question, and explain how your visualization answers the question.
```{r include=FALSE}
library("scales")
```
```{r echo=TRUE}
lego_sales %>% 
  count(theme, age_group) %>%
  group_by(theme) %>%
  mutate(proc = n / sum(n)) %>%
  ggplot(aes(y =theme, x = proc, fill = age_group)) + 
  geom_col(position = 'dodge') + 
  theme_minimal() +
  labs(
    title = "Wich age group is most commonly to buy specific theme of Lego?",
    subtitle = "Lego sales in 2018 for a sample of customers who bought Legos in the US.",
    x = NULL, y = NULL, fill = "Age group"
  ) +
  scale_x_continuous(label = label_percent())

```
